---
title: "PKD Kinta COVID19 Cases for May 2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This page created to make simple visualization for COVID-19 cases specifically for PKD Kinta.

Data for COVID-19 cases for Malaysia were forked from Dr Wan Ariffin github, while the count for COVID-19 cases for PKD Kinta were manually build from daily reported cases in MOH website.

The original reason for this visualization is for prepare manuscript for research articles on EMCO in PKD Kinta.

# Library and Data

```{r, message=FALSE}
library(pacman)
p_load(tidyverse, lubridate,zoo)

covid_19_my <- read_csv("covid-19_my.csv")
covid_19_my_state <- read_csv("covid-19_my_state.csv")
covid_19_kinta <- read_csv("covid-19_kinta.csv")
covid_19_kinta_epidweek <- read_csv("covid-19_kinta_epidweekcases.csv")
covid_19_kinta_mukim_epidweek <- read_csv("covid-19_kinta_mukim_epidweekcases.csv") |>
  pivot_longer(!ME, names_to = "MUKIM", values_to = "IR")

```


# Malaysia & Perak Overview

Malaysia daily new cases, from January 2021, until June 2021

```{r}
covid_19_my |> 
  filter(date > "2021-01-01" & date < "2021-07-01") |> 
  ggplot(aes(x=date, y = new_cases)) + 
  geom_line()

covid_19_my_state |> 
  filter(date > "2021-01-01" & date < "2021-07-01") |> 
  ggplot(aes(x=date, y = new_cases, colour = state)) + 
  geom_line()
```

Perak

```{r}
covid_19_my_state |> 
  filter(state == "PERAK") |> 
  filter(date > "2021-01-01" & date < "2021-07-01") |> 
  ggplot(aes(x=date, y = new_cases)) + 
  geom_line()
```

```{r}
covid_19_my_state |> 
  filter(state == "PERAK") |> 
  filter(date > "2021-01-01" & date < "2021-07-01") |> 
  mutate(newcase7ave = rollmean(new_cases,7, fill = NA)) |> 
  ggplot(aes(x=date, y = newcase7ave)) + 
  geom_line() +
  scale_x_date(date_breaks = "1 months", date_labels = "%F")
```

Kinta

```{r}
covid_19_kinta |> 
  ggplot(aes(x=date, y = new_cases)) + 
  geom_line()
```

```{r}
covid_19_kinta |> 
  mutate(newcase7ave = rollmean(new_cases,7, fill = NA)) |> 
  ggplot(aes(x=date, y = newcase7ave)) + 
  geom_line() +
  scale_x_date(date_breaks = "1 months", date_labels = "%F")
```

# PKD Kinta COVID-19 Cases in 2021 By Epid Week 

Epid Week 1 (ME1) = 3rd - 9th Jan 2021
ME25 = 20th - 26th June 2021

```{r}
covid_19_kinta_epidweek|> 
  ggplot(aes(x=ME, y = new_cases)) + 
  geom_bar(stat="identity") +
  scale_x_continuous(breaks = seq(1,25,1))
```

Incidence Rate per 100,000 population

Daerah Kinta

```{r}
covid_19_kinta_epidweek |> 
  ggplot(aes(x=ME, y=IR)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1,25,1))
```

Mukim in Daerah Kinta

```{r}
covid_19_kinta_mukim_epidweek |> 
  ggplot(aes(x=ME, y=IR, color = MUKIM)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1,25,1))
```

```{r}
ggplot() +
  geom_line(data=covid_19_kinta_epidweek, aes(ME, IR)) +
  geom_line(data=covid_19_kinta_mukim_epidweek, aes(ME, IR, color = MUKIM)) +
#  scale_y_continuous(limits = c(0,90)) +
  scale_x_continuous(breaks = seq(1,25,1)) 
  
```

